---
title: "Age at First Marriage for Women"
author: "Jerome Vonk"
date: "25th February 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The data

This data was extracted from <https://www.gapminder.org/data/>. The continent in which each country resided was added manually by me.
```{r echo=FALSE, include=FALSE}
library(ggplot2)
marriage <- read.csv('age_marriage.csv')
```


## Histogram of Age at First Marriage for Women in 2005
```{r echo=FALSE}
summary(marriage$X2005)
qplot(x = X2005, 
      data = subset(marriage, !is.na(marriage$X2005)),
      binwidth = 1,
      ylab = "Number of occurences",
      xlab = "Age of first marriage for women (2005)",
      color = I('black'), 
      fill = I('#0010F1') )  +
  scale_x_continuous(breaks = seq(16,34,1), limits=c(16,34))
```

## Faceting by continent

This is what we encounter by faceting for continents:

```{r, echo=FALSE, message=FALSE}
qplot(x = X2005, 
      data = subset(marriage, !is.na(marriage$X2005)),
      binwidth = 1,
      ylab = "Number of occurences",
      xlab = "Age of first marriage for women (2005)",
      color = I('black'), 
      fill = I('#0010F1') )  +
  scale_x_continuous(breaks = seq(20,30,5), limits=c(16,34)) +
  facet_wrap(~continent, scales = "free_y")
```

The peak for Africa and Europe look quite different, right? Let's see it numerically:
```{r}
by(marriage$X2005, marriage$continent, summary)
```

Indeed, we find out that Europe shows the highest age at first marriage median (~27.7 years) and Africa the lowest (~21.3). 
